State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-06 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
correcting some typo
suggestion for spatial frequency analysis discussion

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} < \frac{1}{2\rho_{\phi}\nu_{co}}
\end{equation}
where $\rho_{phi}$ is an over-sampling parameter to determine the basis function separation. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{co}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{co}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{co}}= \frac{2\boldsymbol\nu_{co}^\top \boldsymbol\nu_{co}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

@@

\begin{figure*}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{Frequency analysis of the. \textbf{a.} The spatial frequency response of a 1-dimensional neural field basis functions project. The symmetry of the field basis functions, connectivity kernel basis functions and sensors allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field.} 
\end{figure*}
Following this, the distance between basis function, $\Delta_{\phi}$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 2.5, 2.63 mm full width at half maximum. \mike{Analysis said 0.44 from Parham's derivation, but 2.5 still overlaps and allows more high frequency details to be represented.}

>>>>

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} < \frac{1}{2\rho_{\phi}\nu_{co}}
\end{equation}
where $\rho_{\phi}$ is an oversampling parameter to determine the basis function separation. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{co}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{co}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{co}}= \frac{2\boldsymbol\nu_{co}^\top \boldsymbol\nu_{co}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

@@

\begin{figure*}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{Frequency analysis of the. \textbf{a.} The spatial frequency response of a 1-dimensional neural field basis functions project. The symmetry of the field basis functions, connectivity kernel basis functions and sensors allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field.} 
\end{figure*}
Following this, the distance between basis function, $\Delta_{\phi}$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 2.5, 2.63 mm full width at half maximum. \mike{Analysis said 0.44 from Parham's derivation, but 2.5 still overlaps and allows more high frequency details to be represented.}\parham{maybe we can argue that this width is for perfect recunstruction up to 0.19 Hz which requires the smaller distance between basis functions and hence more number of basis functions but we choose to go for lower number of basis function and loosing high frequency contents.}
